---
title: "Repository_for_microarray_analysis"
author: "weitinglin66"
date: "2016/11/15"
output: html_document
---
## Usage of package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(affy)
library(GEOquery)
library(hgu133plus2)
library(annotate)
library(limma)
library(affyPLM)
```

## Introduction

This is the repository for the code used in microarray analysis.

```{r }

```

## Input Microarray cel. file
use the affy package to input the data. it's important to give the metadata for the experiment.

```{r input, echo=FALSE}
 data.path  <- file.path("/Users/Weitinglin/Documents/2016 實驗室資料處理/201510 microarray/raw data 20100114/set 8 CLS CLF Sphere")
 experiment.set <- c ( rep ("set8_withfibroblast" , 3 ) , rep ( "set8_withoutfibroblast" , 3 ), rep("set8_sphere", 3) )
 
 #save the phenotype data
 set<- list.files (data.path, pattern=".CEL")
 phenodata.set               <- matrix ( rep ( set, 2) , ncol = 2 )
 phenodata.set      <- as.data.frame ( phenodata.set )
 colnames ( phenodata.set )   <- c ( "Name" , "FileName" )
 phenodata.set$experiment.set <- experiment.set
 write.table(phenodata.set, paste(data.path,"/phenodata_set.txt", sep = ""),quote=F,sep="\t",row.names=F)
 
 # use the ReadAddy to input
 celfile.set <- ReadAffy ( celfile.path = data.path , phenoData = paste(data.path,"/phenodata_set.txt", sep = "") )
```


## Preprocess

the preprocess step clustering can let us make sure the within group clustering analysis, to ensure the expreiement quality and build the overall sense of the data
```{r clustering analysis}
##clustering 
library(latticeExtra)
dd <- dist2(log2(exprs(Exprs.data)+1))
#dist2 is the method from the genefilter
dd <- as.matrix(dist(t(log2(Exprs.data.median+1)), method = "euclidean"))
#euclidean, maximum, manhattan, canberra, binary, mikowski
dd <- dist2(log2(exprs(Exprs.data )+1))
diag(dd) <- 0
dd.row <- as.dendrogram(hclust(as.dist(dd)))
row.ord <- order.dendrogram(dd.row)
legend <- list(top = list(fun= dendrogramGrob, args = list(x = dd.row, side="top")))
lp <- levelplot(dd[row.ord, row.ord], scales = list(x = list(rot=90)), xlab="", ylab="",legend=legend)
plot(lp)

```


There are plenty way to preprocess the microarray data, such as RMA, MAS5...., and the package Affy(expresso), AffyPLM(three step) provide several way to do the preprocess. 
The overall preprocess method for the affymetrix plateform microarray data include background correction, pm correction, summarization.



```{r preprocess function}
#use the MAS method in one step
Exprs.data <- mas5 ( celfile.set , normalize = FALSE, analysis = "absolute", sc = 500)

#use the expresso to do one step, which have partial flexibility
Exprs.data <- expresso( celfile.set,
                        bg.correct = TRUE,
                        bgcorrect.method = "mas",
                        normalize = FALSE,
                        pmcorrect.method =  "subtractmm",
                        summary.method = "mas")
#pm correction method : "mas" "methods" "pmonly" "subtractmm"
#summary.method "avgdiff" "liwong" "mas" "medianpolish" "playerout"

Exprs.data <- computeExprSet(celfile.set, pmcorrect.method = "pmonly", summary.method = "avgdiff" )
Exprs.data <- generateExprVal.method.avgdiff(exprs(celfile.set))

#combine the background correction, three step to provide more flexibility in microarray preprocession
Exprs.data <- bg.correct(celfile.set, method = "mas") 
#Background correction method:

Exprs.data <- threestep(celfile.set, background = FALSE, normalize = FALSE, summary.method = "median.polish")
#Summarization choice:median.polish tukey.biweight average.log rlm log.average log.median median.log log.2nd.largest lm



```



## Normalization
the quantile normalization method: first quantile normalization ( choose by median or mean )and then log transformation(log2)
```{r quantile_normalization function}
#Quantile Normalization
quantile_normalization <- function(ecelfile.set, method="mean"){
  a <- ecelfile.set
  asort <- apply(a,2,sort)
  loc<-apply(a,2,order)  #原本probe的位置apply
  #step 2 :Takes the means of across rows
  if (method =="mean"){
  amean1<-matrix(rep(apply(asort,1,mean),dim(a)[2]),nrow=length(apply(asort,1,mean)))
  }else if(method == "median"){
    amean1<-matrix(rep(apply(asort,1,median),dim(a)[2]),nrow=length(apply(asort,1,median)))
  }
  #step 3 :Rearranging each column to make the same ordering as the original data
  norm<-matrix(0,dim(a)[1],dim(a)[2])
  
  for(i in 1:dim(amean1)[2]){
    #norm<-matrix(0,dim(amean)[1],1)
    for (j in 1:dim(amean1)[1]){
      #cat(loc[j,i],"\t",i,"\t",j,"\t",amean[j,i],"\n")
      norm[loc[j,i],i]<-amean1[j,i]
    }
  }
  norm <- log2(norm + 1)
  row.names(norm) <- row.names ( ecelfile.set) 
  colnames(norm)  <- colnames ( ecelfile.set)
  return(norm)
}

```


## Prefiltering before the analysis
According to the paper, which indicate the increase power with the prefilter method
```{r prefilter method}

prefilter_choose <- function(eSet, method, parameters){
  if ( method == "filterbyvariance"){
    # print("I'm in filter")
    # print(parameters)
    return(nsFilter(eSet, remove.dupEntrez = FALSE, var.cutoff = parameters[1])$eset)
  } else if ( method == "filterbyexpressionlevel"){
    # print(parameters)
    f1 <- pOverA(parameters[1], parameters[2]) #pOverA(1/3, log2(100))
    f2 <- function(x)(diff(range(x, na.rm = T)) > log2(1.5))
    ff <- filterfun(f1, f2)
    index <- genefilter(exprs(eSet), ff)
    return(eSet[index, ])
  } else if (method == "nofilter"){
    return(eSet)
  }
}

```



## Middle result input

```{r}

```



## Function for DE
The testing_choose, which contain the limma with bayes and wilcoxon rank-sum test with bayes estimation
```{r testing_choose function}
testing_choose <- function(eSet, method, parameters, group){
    CLF_CLS.level    <- gl(n = 2, k =3, labels = c("CLF", "CLS"))
    CLF_sphere.level <- gl(n = 2, k =3, labels = c("CLF", "Sphere"))
    CLS_sphere.level <- gl(n = 2, k =3, labels = c("CLS", "Sphere"))
    paired.level.list <- list(CLF_CLS.level, CLF_sphere.level, CLS_sphere.level)
    contrast.vector <- c("CLF - CLS", "CLF - Sphere", "CLS - Sphere")
    print(paste("Doing the group", group ))
   if (method == "limma") {
     print("Choose the method limma + bayes")
     design.list <- model.matrix(~0+paired.level.list[[group]])
     colnames(design.list) <- levels( paired.level.list[[group]] )
     print("HERE2")
     fit.list <- lmFit(eSet, design.list)
     print(paste("Finish the limma fit in group",group))
    contrast.list <- makeContrasts(contrasts = contrast.vector[group], levels = design.list)
    fit.list <- contrasts.fit(fit.list, contrast.list)
    fit.list <- eBayes(fit.list)
    print(paste("Finish the Bayes in group",group))
    top.list <- topTable(fit.list, coef=contrast.vector[group], number=nrow(fit.list), adjust.method = "BH", p.value = parameters[1]) #parameter=c(0.05)
    print(paste("Finish the testing in group",group))
    return(top.list)
  } else if ( method == "wilcox") {
    print("Choose the method wilcox rank-sum test + bayes")
    print(parameters)
    mw.list <- MultiWilcoxonTest(eSet, paired.level.list[[group]])
    mw <- mw.list
    print("HERE IN WILCOX")
    tidy.data <- data.frame(probe = names(mw@rank.sum.statistics), rank.sum.statistics = unname(mw@rank.sum.statistics))
    rownames(tidy.data) <- names(mw@rank.sum.statistics[])
    tidy.data <- tidy.data[unname(selectSignificant(mw, prior = parameters[1], significance = parameters[2])),]  # parameters = c(0.1, 0.05)
    print(head(tidy.data))
    return(tidy.data)
  }
}

```




the wilcoxon rank-sum test with the parellel computing method boosting
```{r testing_wilcox}
testing_wilcox <- function(eSet, adj.method="BH"){
  registerDoParallel(cores=4)
  expression <- exprs(eSet)
  
  wilcox.result  <- foreach(i = 1:nrow(expression)) %dopar% {
    wilcox.test(expression[i,1:3],expression[i,4:6], correct = FALSE, paired = FALSE) %>%
      tidy %>% mutate(gene = rownames(expression)[i])
  }
  wilcox.result <- invoke(rbind, map(wilcox.result,data.frame))
  #do.call(rbind, lapply(wilcox.result,data.frame))
  wilcox.result <- as.data.frame(wilcox.result)
  rownames(wilcox.result) <- wilcox.result$gene
  wilcox.result <- wilcox.result %>%
    dplyr::select(gene,statistic,p.value,-method, -alternative)%>%
    mutate(adjusted.p = p.adjust(p.value, method=adj.method),
                                            adjusted.method = adj.method)
  return(wilcox.result)
}

```



## Visualization


```{r venndiagram}

area.1 <- length(CLF_CLS.probe)
area.2 <- length(CLF_sphere.probe)
area.3 <- length(CLS_sphere.probe)
area.12 <- length(intersect(CLF_CLS.probe, CLF_sphere.probe))
area.23 <- length(intersect(CLF_sphere.probe,CLS_sphere.probe))
area.13 <- length(intersect(CLF_CLS.probe, CLS_sphere.probe))
area.123 <- length(intersect(intersect(CLF_CLS.probe, CLF_sphere.probe),CLS_sphere.probe))
area.list <- list(area.1, area.2, area.3, area.12, area.23, area.13, area.123)


#make the Venn Diagram

library(VennDiagram)
#save the output vennDiagram into .tiff file
venn.diagram(list(CLF_CLS=CLF_CLS.probe , CLS_Sphere=CLS_sphere.probe, CLF_Sphere=CLF_sphere.probe), fill=c("red","green","yellow"), alpha=c(0.5, 0.5, 0.5), filename="VennDiagram_NEW_CLF-CLS_CLS-Sphere_CLF-Sphere.tiff")

#plot into the device
draw.triple.venn(area1 = area.list[[1]], area2 = area.list[[2]], area3 = area.list[[3]], n12 = area.list[[4]], n23 = area.list[[5]], n13 = area.list[[6]],
                   n123 = area.list[[7]], category = c("CLF_CLS", "CLF_Sphere", "CLS_Sphere"), lty = "blank",
                   fill = c("skyblue", "pink1", "mediumorchid"))

```











